## Permanovas

#### Load required data

```{r loaddata_2}
meta <- column_to_rownames(sample_metadata, "Tube_code")
```

### 1. Are the wild populations similar?

#### Wild: P.muralis vs P.liolepis

```{r beta_div_neutral_wild_, echo=TRUE, results = FALSE, message=FALSE, warning=FALSE, comments=FALSE}
wild <- meta %>%
  filter(time_point == "0_Wild")

# Create a temporary modified version of genome_counts_filt
temp_genome_counts <- transform(genome_counts_filt, row.names = genome_counts_filt$genome)
temp_genome_counts$genome <- NULL

wild.counts <- temp_genome_counts[, which(colnames(temp_genome_counts) %in% rownames(wild))]
identical(sort(colnames(wild.counts)), sort(as.character(rownames(wild))))

wild_nmds <- sample_metadata %>%
  filter(time_point == "0_Wild")
```

#### Number of samples used

```{r beta_div_neutral_wild_samples, echo=FALSE, results = TRUE, message=FALSE, warning=FALSE, comments=FALSE}
nrow(wild)
```

```{r beta_div_neutral_wild_1, echo=TRUE, results = FALSE, message=FALSE, warning=FALSE, comments=FALSE, eval=FALSE}
beta_div_richness_wild<-hillpair(data=wild.counts, q=0)
beta_div_neutral_wild<-hillpair(data=wild.counts, q=1)
beta_div_phylo_wild<-hillpair(data=wild.counts, q=1, tree=genome_tree)
beta_div_func_wild<-hillpair(data=wild.counts, q=1, dist=dist)
```

#### Richness

```{r beta_richness_permanova_1, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_richness_wild$S, wild$species) %>% permutest(., pairwise = TRUE)
```

```{r beta_richness_permanova, echo=FALSE, warning=FALSE, comments="", message=FALSE, results=TRUE}
adonis2(formula=beta_div_richness_wild$S ~ species, data=wild[labels(beta_div_richness_wild$S),], permutations=999) %>%
			as.matrix() %>%
			kable()
```


#### Neutral

```{r beta_neutral_permanova_1, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_neutral_wild$S, wild$species) %>% permutest(., pairwise = TRUE)
```

```{r beta_neutral_permanova, echo=FALSE, warning=FALSE, comments="", message=FALSE, results=TRUE}
adonis2(formula=beta_div_neutral_wild$S ~ species, data=wild[labels(beta_div_neutral_wild$S),], permutations=999) %>%
			as.matrix() %>%
			kable()
```

#### Phylogenetic

```{r beta_phylo_permanova_1, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_phylo_wild$S, wild$species) %>% permutest(., pairwise = TRUE)
```

```{r beta_phylo_permanova, echo=FALSE, warning=FALSE, comments="", message=FALSE, results=TRUE}
adonis2(formula=beta_div_phylo_wild$S ~ species, data=wild[labels(beta_div_phylo_wild$S),], permutations=999) %>%
			as.matrix() %>%
			kable()
```

#### Functional

```{r beta_func_permanova_!, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_func_wild$S, wild$species) %>% permutest(., pairwise = TRUE)
```

```{r beta_func_permanova, echo=FALSE, warning=FALSE, comments="", message=FALSE, results=TRUE}
adonis2(formula=beta_div_func_wild$S ~ species, data=wild[labels(beta_div_func_wild$S),], permutations=999) %>%
			as.matrix() %>%
			kable()
```

```{r beta_neutral_nmds, echo=TRUE, warning=FALSE, comments="", message=FALSE, results="hide"}
beta_q0n_nmds_wild <- beta_div_richness_wild$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE, trace=FALSE) %>%
				vegan::scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(wild_nmds, by = join_by(sample == Tube_code))

beta_q1n_nmds_wild <- beta_div_neutral_wild$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE, trace=FALSE) %>%
				vegan::scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(wild_nmds, by = join_by(sample == Tube_code))

beta_q1p_nmds_wild <- beta_div_phylo_wild$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				vegan::scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(wild_nmds, by = join_by(sample == Tube_code))

beta_q1f_nmds_wild <- beta_div_func_wild$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				vegan::scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(wild_nmds, by = join_by(sample == Tube_code))
```

```{r beta_neutral_nmds_plot, echo=FALSE, results=TRUE}
p0<-beta_q0n_nmds_wild %>%
			group_by(Population) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=Population)) +
				scale_color_manual(name="Population",
                    breaks=c("Cold_wet","Hot_dry"),
                    labels=c("Cold","Hot"),
                    values=c('#008080',"#d57d2c")) +
        scale_shape_manual(name="Type",
                           breaks=c("Control", "Hot_control", "Treatment"),
                           labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                         values=c("circle","square","triangle")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Richness beta diversity") +
				theme_classic() +
				theme(legend.position="none")

p1<-beta_q1n_nmds_wild %>%
			group_by(Population) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=Population, shape=type)) +
				scale_color_manual(name="Population",
                    breaks=c("Cold_wet","Hot_dry"),
                    labels=c("Cold","Hot"),
                    values=c('#008080',"#d57d2c")) +
        scale_shape_manual(name="Type",
                           breaks=c("Control", "Hot_control", "Treatment"),
                           labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                         values=c("circle","square","triangle")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Neutral beta diversity") +
				theme_classic() +
				theme(legend.position="none")

p2<-beta_q1p_nmds_wild %>%
			group_by(Population) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=Population)) +
				scale_color_manual(name="Population",
                    breaks=c("Cold_wet","Hot_dry"),
                    labels=c("Cold","Hot"),
                    values=c('#008080',"#d57d2c")) +
        scale_shape_manual(name="Type",
                           breaks=c("Control", "Hot_control", "Treatment"),
                           labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                         values=c("circle","square","triangle")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (),x="NMDS1 \n Phylogenetic beta diversity") +
				theme_classic() +
				theme(legend.position="none") 

p3<-beta_q1f_nmds_wild %>%
			group_by(Population) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=Population, label=sample)) +
				scale_color_manual(name="Population",
                    breaks=c("Cold_wet","Hot_dry"),
                    labels=c("Cold","Hot"),
                    values=c('#008080',"#d57d2c")) +
        scale_shape_manual(name="Type",
                           breaks=c("Control", "Hot_control", "Treatment"),
                           labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                         values=c("circle","square","triangle")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (), x="NMDS1 \n Functional beta diversity") +
				theme_classic() +
				theme(legend.position="none") +
				guides(color=guide_legend(title="Species"))
```

```{r beta_neutral_nmds_plot_final, echo=FALSE, results=TRUE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
ggarrange(p0, p1, p2, p3, ncol=2, nrow=2, common.legend = TRUE, legend="right")
```

### 2. Effect of acclimation

```{r beta_div_neutral_accli, echo=TRUE, results = FALSE, message=FALSE, warning=FALSE, comments=FALSE}
accli <- meta %>%
  filter(time_point == "1_Acclimation")

# Create a temporary modified version of genome_counts_filt
temp_genome_counts <- transform(genome_counts_filt, row.names = genome_counts_filt$genome)
temp_genome_counts$genome <- NULL

accli.counts <- temp_genome_counts[, which(colnames(temp_genome_counts) %in% rownames(accli))]
identical(sort(colnames(accli.counts)), sort(as.character(rownames(accli))))

accli_nmds <- sample_metadata %>%
  filter(time_point == "1_Acclimation")
```

#### Number of samples used

```{r beta_div_neutral_accli_samples, echo=FALSE, results = TRUE, message=FALSE, warning=FALSE, comments=FALSE}
nrow(accli)
```

```{r beta_div_neutral_accli_1, echo=TRUE, results = FALSE, message=FALSE, warning=FALSE, comments=FALSE, eval=FALSE}
beta_div_richness_accli<-hillpair(data=accli.counts, q=0)
beta_div_neutral_accli<-hillpair(data=accli.counts, q=1)
beta_div_phylo_accli<-hillpair(data=accli.counts, q=1, tree=genome_tree)
beta_div_func_accli<-hillpair(data=accli.counts, q=1, dist=dist)
```

#### Richness

```{r beta_richness_permanova_1_accli, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_richness_accli$S, accli$Population) %>% permutest(., pairwise = TRUE)
```

```{r beta_richness_permanova_accli, echo=FALSE, warning=FALSE, comments="", message=FALSE, results=TRUE}
adonis2(formula=beta_div_richness_accli$S ~ Population, data=accli[labels(beta_div_richness_accli$S),], permutations=999) %>%
  as.matrix() %>%
  kable()
```


#### Neutral

```{r beta_neutral_permanova_1_accli, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_neutral_accli$S, accli$Population) %>% permutest(., pairwise = TRUE)
```

```{r beta_neutral_permanova_accli, echo=FALSE, warning=FALSE, comments="", message=FALSE, results=TRUE}
adonis2(formula=beta_div_neutral_accli$S ~ Population, data=accli[labels(beta_div_neutral_accli$S),], permutations=999) %>%
  as.matrix() %>%
  kable()
```

#### Phylogenetic

```{r beta_phylo_permanova_1_accli, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_phylo_accli$S, accli$Population) %>% permutest(., pairwise = TRUE)
```

```{r beta_phylo_permanova_accli, echo=FALSE, warning=FALSE, comments="", message=FALSE, results=TRUE}
adonis2(formula=beta_div_phylo_accli$S ~ Population, data=accli[labels(beta_div_phylo_accli$S),], permutations=999) %>%
  as.matrix() %>%
  kable()
```

#### Functional

```{r beta_func_permanova_accli_1, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_func_accli$S, accli$Population) %>% permutest(., pairwise = TRUE)
```

```{r beta_func_permanova_accli, echo=FALSE, warning=FALSE, comments="", message=FALSE, results=TRUE}
adonis2(formula=beta_div_func_accli$S ~ Population, data=accli[labels(beta_div_func_accli$S),], permutations=999) %>%
  as.matrix() %>%
  kable()
```


```{r beta_neutral_nmds_accli, echo=TRUE, warning=FALSE, comments="", message=FALSE, results="hide"}
beta_q0n_nmds_accli <- beta_div_richness_accli$S %>%
  metaMDS(.,trymax = 500, k=2, verbosity=FALSE, trace=FALSE) %>%
  vegan::scores() %>%
  as_tibble(., rownames = "sample") %>%
  left_join(accli_nmds, by = join_by(sample == Tube_code))

beta_q1n_nmds_accli <- beta_div_neutral_accli$S %>%
  metaMDS(.,trymax = 500, k=2, verbosity=FALSE, trace=FALSE) %>%
  vegan::scores() %>%
  as_tibble(., rownames = "sample") %>%
  left_join(accli_nmds, by = join_by(sample == Tube_code))

beta_q1p_nmds_accli <- beta_div_phylo_accli$S %>%
  metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
  vegan::scores() %>%
  as_tibble(., rownames = "sample") %>%
  left_join(accli_nmds, by = join_by(sample == Tube_code))

beta_q1f_nmds_accli <- beta_div_func_accli$S %>%
  metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
  vegan::scores() %>%
  as_tibble(., rownames = "sample") %>%
  left_join(accli_nmds, by = join_by(sample == Tube_code))
```

```{r beta_neutral_nmds_plot_accli, echo=FALSE, results=TRUE}
p0<-beta_q0n_nmds_accli %>%
  group_by(Population) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(., aes(x=NMDS1,y=NMDS2, color=Population)) +
  scale_color_manual(name="Population",
                     breaks=c("Cold_wet","Hot_dry"),
                     labels=c("Cold","Hot"),
                     values=c('#008080',"#d57d2c")) +
  scale_shape_manual(name="Type",
                     breaks=c("Control", "Hot_control", "Treatment"),
                     labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                     values=c("circle","square","triangle")) +
  geom_point(size=2) +
  geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
  labs(y = "NMDS2", x="NMDS1 \n Richness beta diversity") +
  theme_classic() +
  theme(legend.position="none")

p1<-beta_q1n_nmds_accli %>%
  group_by(Population) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(., aes(x=NMDS1,y=NMDS2, color=Population, shape=type)) +
  scale_color_manual(name="Population",
                     breaks=c("Cold_wet","Hot_dry"),
                     labels=c("Cold","Hot"),
                     values=c('#008080',"#d57d2c")) +
  scale_shape_manual(name="Type",
                     breaks=c("Control", "Hot_control", "Treatment"),
                     labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                     values=c("circle","square","triangle")) +
  geom_point(size=2) +
  geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
  labs(y = "NMDS2", x="NMDS1 \n Neutral beta diversity") +
  theme_classic() +
  theme(legend.position="none")

p2<-beta_q1p_nmds_accli %>%
  group_by(Population) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(., aes(x=NMDS1,y=NMDS2, color=Population)) +
  scale_color_manual(name="Population",
                     breaks=c("Cold_wet","Hot_dry"),
                     labels=c("Cold","Hot"),
                     values=c('#008080',"#d57d2c")) +
  scale_shape_manual(name="Type",
                     breaks=c("Control", "Hot_control", "Treatment"),
                     labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                     values=c("circle","square","triangle")) +
  geom_point(size=2) +
  geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
  labs(y= element_blank (),x="NMDS1 \n Phylogenetic beta diversity") +
  theme_classic() +
  theme(legend.position="none") 

p3<-beta_q1f_nmds_accli %>%
  group_by(Population) %>%
  mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
  mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(., aes(x=NMDS1,y=NMDS2, color=Population, label=sample)) +
  scale_color_manual(name="Population",
                     breaks=c("Cold_wet","Hot_dry"),
                     labels=c("Cold","Hot"),
                     values=c('#008080',"#d57d2c")) +
  scale_shape_manual(name="Type",
                     breaks=c("Control", "Hot_control", "Treatment"),
                     labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                     values=c("circle","square","triangle")) +
  geom_point(size=2) +
  geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
  labs(y= element_blank (), x="NMDS1 \n Functional beta diversity") +
  theme_classic() +
  theme(legend.position="none") +
  guides(color=guide_legend(title="Species"))
```

```{r beta_neutral_nmds_plot_final_accli, echo=FALSE, results=TRUE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
ggarrange(p0, p1, p2, p3, ncol=2, nrow=2, common.legend = TRUE, legend="right")
```

### 3. Comparison between Wild and Acclimation

```{r beta_div_neutral_accli1, echo=TRUE, results=FALSE,message=FALSE, warning=FALSE, comments=FALSE}
accli1 <- meta  %>%
  filter(time_point == "0_Wild" | time_point == "1_Acclimation")

temp_genome_counts <- transform(genome_counts_filt, row.names = genome_counts_filt$genome)
temp_genome_counts$genome <- NULL

accli1.counts <- temp_genome_counts[,which(colnames(temp_genome_counts) %in% rownames(accli1))]
identical(sort(colnames(accli1.counts)),sort(as.character(rownames(accli1))))

accli1_nmds <- sample_metadata %>%
  filter(time_point == "0_Wild" | time_point == "1_Acclimation")
```

#### Number of samples used

```{r beta_div_neutral_accli1_samples, echo=FALSE, results = TRUE, message=FALSE, warning=FALSE, comments=FALSE}
nrow(accli1)
```

```{r beta_div_neutral_accli1_1, echo=TRUE, results=FALSE,message=FALSE, warning=FALSE, comments=FALSE, eval=FALSE}
beta_div_richness_accli1<-hillpair(data=accli1.counts, q=0)
beta_div_neutral_accli1<-hillpair(data=accli1.counts, q=1)
beta_div_phylo_accli1<-hillpair(data=accli1.counts, q=1, tree=genome_tree)
beta_div_func_accli1<-hillpair(data=accli1.counts, q=1, dist=dist)
```

##### Richness

```{r beta_div_richness_accli1_5, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_richness_accli1$S, accli1$time_point) %>% permutest(., pairwise = TRUE)
```

```{r beta_div_richness_accli1_4, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_richness_accli1$S ~ time_point*species, data=accli1[labels(beta_div_neutral_accli1$S),], permutations=999, strata=accli1$individual) %>%
		as.matrix() %>%
		kable()
```

##### Neutral

```{r beta_div_neutral_accli1_5, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_neutral_accli1$S, accli1$time_point) %>% permutest(., pairwise = TRUE)
```

```{r beta_div_neutral_accli1_4, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_neutral_accli1$S ~ time_point*species, data=accli1[labels(beta_div_neutral_accli1$S),], permutations=999, strata=accli1$individual) %>%
		as.matrix() %>%
		kable()
```

##### Phylogenetic

```{r beta_div_phylo_accli1_5, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_phylo_accli1$S, accli1$time_point) %>% permutest(., pairwise = TRUE)
```

```{r beta_div_phylo_accli1_4, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_phylo_accli1$S ~ time_point*species, data=accli1[labels(beta_div_phylo_accli1$S),], permutations=999, strata=accli1$individual) %>%
		as.matrix() %>%
		kable()
```

##### Functional

```{r beta_div_func_accli1_5, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_func_accli1$S, accli1$time_point) %>% permutest(., pairwise = TRUE)
```

```{r beta_div_func_accli1_4, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_func_accli1$S ~ time_point*species, data=accli1[labels(beta_div_func_accli1$S),], permutations=999, strata=accli1$individual) %>%
		as.matrix() %>%
		kable()
```


```{r beta_neutral_nmds_accli1, echo=TRUE, warning=FALSE, comments="", message=FALSE, results="hide"}
beta_richness_nmds_accli1 <- beta_div_richness_accli1$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(accli1_nmds, by = c("sample" = "Tube_code"))

beta_neutral_nmds_accli1 <- beta_div_neutral_accli1$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(accli1_nmds, by = c("sample" = "Tube_code"))

beta_phylo_nmds_accli1 <- beta_div_phylo_accli1$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(accli1_nmds, by = join_by(sample == Tube_code))

beta_func_nmds_accli1 <- beta_div_func_accli1$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(accli1_nmds, by = join_by(sample == Tube_code))
```

```{r beta_neutral_nmds_accli1_plot, echo=FALSE, results=FALSE}
p0<-beta_richness_nmds_accli1 %>%
			group_by(individual) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type, shape=time_point)) +
        scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "accli1ment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Richness beta diversity") +
				theme_classic() +
				theme(legend.position="none")

p1<-beta_neutral_nmds_accli1 %>%
			group_by(individual) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type, shape=time_point)) +
        scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "accli1ment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Neutral beta diversity") +
				theme_classic() +
				theme(legend.position="none")
  
p2<-beta_phylo_nmds_accli1 %>%
			group_by(individual) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type, shape=time_point)) +
				scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "accli1ment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (), x="NMDS1 \n Phylogenetic beta diversity") +
				theme_classic() +
				theme(legend.position="none")

p3<-beta_func_nmds_accli1 %>%
			group_by(individual) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type, shape=time_point)) +
				scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "accli1ment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (), x="NMDS1 \n Functional beta diversity") +
				theme_classic()+
				theme(legend.position="none")
```

```{r beta_neutral_nmds_plot_accli1_final, echo=FALSE, results=TRUE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
ggarrange(p0, p1, p2, p3, ncol=2, nrow=2, common.legend = TRUE, legend="right")
```

### 4. Do the antibiotics work?

#### Acclimation vs antibiotics

```{r beta_div_neutral_treat, echo=TRUE, results=FALSE,message=FALSE, warning=FALSE, comments=FALSE}
treat <- meta  %>%
  filter(time_point == "1_Acclimation" | time_point == "2_Antibiotics")

temp_genome_counts <- transform(genome_counts_filt, row.names = genome_counts_filt$genome)
temp_genome_counts$genome <- NULL

treat.counts <- temp_genome_counts[,which(colnames(temp_genome_counts) %in% rownames(treat))]
identical(sort(colnames(treat.counts)),sort(as.character(rownames(treat))))

treat_nmds <- sample_metadata %>%
  filter(time_point == "1_Acclimation" | time_point == "2_Antibiotics")
```

#### Number of samples used

```{r beta_div_neutral_treat_samples, echo=FALSE, results = TRUE, message=FALSE, warning=FALSE, comments=FALSE}
nrow(treat)
```

```{r beta_div_neutral_treat_1, echo=TRUE, results=FALSE,message=FALSE, warning=FALSE, comments=FALSE, eval=FALSE}
beta_div_richness_treat<-hillpair(data=treat.counts, q=0)
beta_div_neutral_treat<-hillpair(data=treat.counts, q=1)
beta_div_phylo_treat<-hillpair(data=treat.counts, q=1, tree=genome_tree)
beta_div_func_treat<-hillpair(data=treat.counts, q=1, dist=dist)
```

##### Richness

```{r beta_div_richness_treat_5, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_richness_treat$S, treat$time_point) %>% permutest(., pairwise = TRUE)
```

```{r beta_div_richness_treat_4, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_richness_treat$S ~ time_point+species+individual, data=treat[labels(beta_div_neutral_treat$S),], permutations=999) %>%
		as.matrix() %>%
		kable()
```

##### Neutral

```{r beta_div_neutral_treat_5, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_neutral_treat$S, treat$time_point) %>% permutest(., pairwise = TRUE)
```

```{r beta_div_neutral_treat_4, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_neutral_treat$S ~ time_point+species+individual, data=treat[labels(beta_div_neutral_treat$S),], permutations=999) %>%
		as.matrix() %>%
		kable()
```


##### Phylogenetic

```{r beta_div_phylo_treat_5, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_phylo_treat$S, treat$time_point) %>% permutest(., pairwise = TRUE)
```

```{r beta_div_phylo_treat_4, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_phylo_treat$S ~ time_point+species+individual, data=treat[labels(beta_div_phylo_treat$S),], permutations=999) %>%
		as.matrix() %>%
		kable()
```

##### Functional

```{r beta_div_func_treat_5, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_func_treat$S, treat$time_point) %>% permutest(., pairwise = TRUE)
```

```{r beta_div_func_treat_4, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_func_treat$S ~ time_point+species+individual, data=treat[labels(beta_div_func_treat$S),], permutations=999) %>%
		as.matrix() %>%
		kable()
```


```{r beta_neutral_nmds_treat, echo=TRUE, warning=FALSE, comments="", message=FALSE, results="hide"}
beta_richness_nmds_treat <- beta_div_richness_treat$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(treat_nmds, by = c("sample" = "Tube_code"))

beta_neutral_nmds_treat <- beta_div_neutral_treat$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(treat_nmds, by = c("sample" = "Tube_code"))

beta_phylo_nmds_treat <- beta_div_phylo_treat$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(treat_nmds, by = join_by(sample == Tube_code))

beta_func_nmds_treat <- beta_div_func_treat$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(treat_nmds, by = join_by(sample == Tube_code))
```

```{r beta_neutral_nmds_treat_plot, echo=FALSE, results=FALSE}
p0<-beta_richness_nmds_treat %>%
			group_by(individual) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type, shape=time_point)) +
        scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Richness beta diversity") +
				theme_classic() +
				theme(legend.position="none")

p1<-beta_neutral_nmds_treat %>%
			group_by(individual) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type, shape=time_point)) +
        scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Neutral beta diversity") +
				theme_classic() +
				theme(legend.position="none")
  
p2<-beta_phylo_nmds_treat %>%
			group_by(individual) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type, shape=time_point)) +
				scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (), x="NMDS1 \n Phylogenetic beta diversity") +
				theme_classic() +
				theme(legend.position="none")

p3<-beta_func_nmds_treat %>%
			group_by(individual) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type, shape=time_point)) +
				scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (), x="NMDS1 \n Functional beta diversity") +
				theme_classic()+
				theme(legend.position="none")
```

```{r beta_neutral_nmds_plot_treat_final, echo=FALSE, results=TRUE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
ggarrange(p0, p1, p2, p3, ncol=2, nrow=2, common.legend = TRUE, legend="right")
```


### 5. Does the FMT work?

#### Comparison between FMT2 vs Post-FMT2

```{r beta_div_neutral_transplant3, echo=TRUE, results=FALSE,message=FALSE, warning=FALSE, comments=FALSE}
#Create newID to identify duplicated samples
transplants_metadata<-sample_metadata%>%
  mutate(Tube_code=str_remove_all(Tube_code, "_a"))
transplants_metadata$newID <- paste(transplants_metadata$Tube_code, "_", transplants_metadata$individual)

transplant3<-transplants_metadata%>%
  filter(time_point == "4_Transplant2" | time_point == "6_Post-FMT2")%>%
  column_to_rownames("newID")

transplant3_nmds <- transplants_metadata %>%
  filter(time_point == "4_Transplant2" | time_point == "6_Post-FMT2")

full_counts<-temp_genome_counts %>%
    t()%>%
    as.data.frame()%>%
    rownames_to_column("Tube_code")%>%
    full_join(transplants_metadata,by = join_by(Tube_code == Tube_code))

transplant3_counts<-full_counts %>%
  filter(time_point == "4_Transplant2" | time_point == "6_Post-FMT2") %>%
  subset(select=-c(315:324)) %>%
  column_to_rownames("newID")%>%
  subset(select=-c(1))%>%
  t() %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric)

identical(sort(colnames(transplant3_counts)),sort(as.character(rownames(transplant3))))
```

#### Number of samples used

```{r beta_div_neutral_treat_samples_1, echo=FALSE, results = TRUE, message=FALSE, warning=FALSE, comments=FALSE}
nrow(transplant3)
```


```{r beta_div_neutral_transplant3_1, echo=TRUE, results=FALSE,message=FALSE, warning=FALSE, comments=FALSE, eval=FALSE}
beta_div_richness_transplant3<-hillpair(data=transplant3_counts, q=0)
beta_div_neutral_transplant3<-hillpair(data=transplant3_counts, q=1)
beta_div_phylo_transplant3<-hillpair(data=transplant3_counts, q=1, tree=genome_tree)
beta_div_func_transplant3<-hillpair(data=transplant3_counts, q=1, dist=dist)
```

```{r beta_div_neutral_transplant3_2, echo=TRUE, results=FALSE,message=FALSE, warning=FALSE, comments=FALSE}
#Arrange of metadata dataframe
transplant3_arrange<-transplant3[labels(beta_div_neutral_transplant3$S),]
```

##### Richness

```{r beta_richness_permanova_transplant3_4, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_richness_transplant3$S ~ species+time_point+type, data=transplant3[labels(beta_div_richness_transplant3$S),], permutations=999, strata=transplant3$individual) %>%
		as.matrix() %>%
		kable()
```

```{r beta_richness_permanova_transplant3_5, echo=FALSE, warning=FALSE, comments="", message=FALSE}
pairwise<-pairwise.adonis(beta_div_richness_transplant3$S,transplant3_arrange$type, perm=999)
pairwise
```

##### Neutral

```{r beta_neutral_permanova_transplant3_4, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_neutral_transplant3$S ~ species+time_point+type, data=transplant3[labels(beta_div_neutral_transplant3$S),], permutations=999, strata=transplant3$individual) %>%
		as.matrix() %>%
		kable()
```

```{r beta_neutral_permanova_transplant3_5, echo=FALSE, warning=FALSE, comments="", message=FALSE}
pairwise<-pairwise.adonis(beta_div_neutral_transplant3$S,transplant3_arrange$type, perm=999)
pairwise
```

##### Phylogenetic

```{r beta_phylo_permanova_transplant3_4, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_phylo_transplant3$S ~ species+time_point+type, data=transplant3[labels(beta_div_phylo_transplant3$S),], permutations=999, strata=transplant3$individual) %>%
		as.matrix() %>%
		kable()
```

```{r beta_phylo_permanova_transplant3_5, echo=FALSE, warning=FALSE, comments="", message=FALSE}
pairwise<-pairwise.adonis(beta_div_phylo_transplant3$S,transplant3_arrange$type, perm=999)
pairwise
```

##### Functional

```{r beta_func_permanova_transplant3_4, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_func_transplant3$S ~ species+time_point+type, data=transplant3[labels(beta_div_func_transplant3$S),], permutations=999, strata=transplant3$individual) %>%
		as.matrix() %>%
		kable()
```

```{r beta_func_permanova_transplant3_5, echo=FALSE, warning=FALSE, comments="", message=FALSE}
pairwise<-pairwise.adonis(beta_div_func_transplant3$S,transplant3_arrange$type, perm=999)
pairwise
```

```{r beta_neutral_nmds_transplant3, echo=TRUE, warning=FALSE, comments="", message=FALSE, results="hide"}
beta_richness_nmds_transplant3 <- beta_div_richness_transplant3$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(transplant3_nmds, by = join_by(sample == newID))

beta_neutral_nmds_transplant3 <- beta_div_neutral_transplant3$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(transplant3_nmds, by = join_by(sample == newID))

beta_phylo_nmds_transplant3 <- beta_div_phylo_transplant3$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(transplant3_nmds, by = join_by(sample == newID))

beta_func_nmds_transplant3 <- beta_div_func_transplant3$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(transplant3_nmds, by = join_by(sample == newID))
```

```{r beta_neutral_nmds_transplant3_plot, eho=FALSE, results=FALSE}
p0<-beta_richness_nmds_transplant3 %>%
			group_by(individual) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type, shape=time_point)) +
        scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Richness beta diversity") +
				theme_classic() +
				theme(legend.position="none")

p1<-beta_neutral_nmds_transplant3 %>%
			group_by(individual) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type, shape=time_point)) +
        scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Neutral beta diversity") +
				theme_classic() +
				theme(legend.position="none")
  
p2<-beta_phylo_nmds_transplant3 %>%
			group_by(individual) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type, shape=time_point)) +
				scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (), x="NMDS1 \n Phylogenetic beta diversity") +
				theme_classic() +
				theme(legend.position="none")

p3<-beta_func_nmds_transplant3 %>%
			group_by(individual) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type, shape=time_point)) +
				scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (), x="NMDS1 \n Functional beta diversity") +
				theme_classic()+
				theme(legend.position="none")
```

```{r beta_neutral_nmds_plot_transplant3_final, echo=FALSE, results=TRUE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
ggarrange(p0, p1, p2, p3, ncol=2, nrow=2, common.legend = TRUE, legend="right")
```



#### Comparison between the different experimental time points
##### Comparison against Acclimation samples with combined Transplant samples

```{r diss_data1_3, echo=FALSE, results=FALSE}
transplants_metadata<-transplants_metadata%>%
  filter(!(Tube_code == "AD45" | Tube_code == "AD48"))

#Create newID to identify duplicated samples
transplant6<-transplants_metadata%>%
  filter(time_point == "3_Transplant1" | time_point == "4_Transplant2" | time_point == "6_Post-FMT2" |time_point == "1_Acclimation")%>%
  column_to_rownames("newID")

#Transform time_point names to combine the transplant samples
transplant6 <- transplant6 %>%
  mutate(time_point = recode(time_point,
                             "3_Transplant1" = "Transplant",
                             "4_Transplant2" = "Transplant"))

#Transform time_point names to combine the transplant samples
transplant6_pairwise <- transplants_metadata %>%
  filter(time_point == "3_Transplant1" | time_point == "4_Transplant2" | time_point == "6_Post-FMT2"|time_point == "1_Acclimation")

transplant6_pairwise <- transplant6_pairwise %>%
  mutate(time_point = recode(time_point,
                             "3_Transplant1" = "Transplant",
                             "4_Transplant2" = "Transplant"))

full_counts_new<-temp_genome_counts %>%
  t()%>%
  as.data.frame()%>%
  rownames_to_column("Tube_code")%>%
  full_join(transplants_metadata,by = join_by(Tube_code == Tube_code))

transplant6_counts<-full_counts %>%
  filter(time_point == "3_Transplant1" | time_point == "4_Transplant2" | time_point == "6_Post-FMT2"|time_point == "1_Acclimation") %>%
  subset(select=-c(315:324)) %>%
  column_to_rownames("newID")%>%
  subset(select=-c(1))%>%
  t() %>%
  as.data.frame() %>%
  mutate_if(is.character, as.numeric)

identical(sort(colnames(transplant6_counts)),sort(as.character(rownames(transplant6))))
```

```{r diss_data1_4, echo=FALSE, results=FALSE}
beta_div_neutral_transplant6<-hillpair(data=transplant6_counts, q=1)

transplant6_beta_div_neutral<-as.matrix(beta_div_neutral_transplant6$S) #convert the pairwise comparisons into a matrix

transplant6_beta_div_neutral_df <- as.data.frame(as.table(as.matrix(transplant6_beta_div_neutral))) #convert the matrix to a data frame with pairwise values

transplant6_beta_div_neutral_df <- transplant6_beta_div_neutral_df %>% #remove the duplicated pairwise comparisons
  filter(Var1 != Var2) %>%
  mutate(pair = pmap_chr(list(Var1, Var2), ~paste(sort(c(..1, ..2)), collapse = "_"))) %>%
  distinct(pair, .keep_all = TRUE) %>%
  select(Sample_1 = Var1, Sample_2 = Var2, Distance = Freq) %>%
  arrange(Sample_1, Sample_2)


transplant6_beta_div_neutral_df <- transplant6_beta_div_neutral_df %>% #keep only the pairwise comparisons between the same individual
  mutate(Sample_1_part = sub("^[^_]*_", "", Sample_1),
         Sample_2_part = sub("^[^_]*_", "", Sample_2)) %>%
  filter(Sample_1_part == Sample_2_part) 
#%>%
#select(Sample_1, Sample_2, Distance)


transplant6_beta_div_neutral_met<-transplant6_beta_div_neutral_df %>% #merge with the metadata associated to the pairwise comparisons for Sample_1
  inner_join(transplant6_pairwise, by = join_by(Sample_1 == "newID"))

# Extract the part before the first `_` for Sample_1 and Sample_2
transplant6_beta_div_neutral_met$Sample_1_part <- sub("_.*", "", transplant6_beta_div_neutral_met$Sample_1)

# Create a named vector for matching Tube_code with Time_point
time_point_map_new <- setNames(transplant6_beta_div_neutral_met$time_point, transplant6_beta_div_neutral_met$Tube_code)

# Function to replace the part before the first _ with the corresponding Time_point
replace_with_time_point_new <- function(sample_part, tube_code, time_point_map_new) {
  if (tube_code %in% names(time_point_map_new)) {
    return(time_point_map_new[tube_code])
  } else {
    return(sample_part)
  }
}

# Apply the function to Sample_1_part
transplant6_beta_div_neutral_met$Sample_1_part <- mapply(replace_with_time_point_new,
                                                         transplant6_beta_div_neutral_met$Sample_1_part,
                                                         transplant6_beta_div_neutral_met$Tube_code,
                                                         list(time_point_map_new))

transplant6_beta_div_neutral_met<-transplant6_beta_div_neutral_met %>%
  subset(select=-c(6:16))

transplant6_beta_div_neutral_met$Sample_2_part <- sub("_.*", "", transplant6_beta_div_neutral_met$Sample_2)


transplant6_beta_div_neutral_met<-transplant6_beta_div_neutral_met %>% #merge with the metadata associated to the pairwise comparisons for Sample_2
  inner_join(transplant6_pairwise, by = join_by(Sample_2 == "newID"))

# Create a named vector again for matching Tube_code with Time_point
time_point_map_new <- setNames(transplant6_beta_div_neutral_met$time_point, transplant6_beta_div_neutral_met$Tube_code)


replace_with_time_point_new <- function(sample_part, tube_code, time_point_map_new) {
  if (tube_code %in% names(time_point_map_new)) {
    return(time_point_map_new[tube_code])
  } else {
    return(sample_part)
  }
}

# Apply the function to Sample_2_part
transplant6_beta_div_neutral_met$Sample_2_part <- mapply(replace_with_time_point_new,
                                                         transplant6_beta_div_neutral_met$Sample_2_part,
                                                         transplant6_beta_div_neutral_met$Tube_code,
                                                         list(time_point_map_new))


# Create a new variable to plot the distances between the time_points
transplant6_beta_div_neutral_met$comparison <- paste(transplant6_beta_div_neutral_met$Sample_1_part, "-",transplant6_beta_div_neutral_met$Sample_2_part)


##plot the differences between acclimation and post-transplant and transplant and post-transplant

# Reorder the 'comparison' factor
transplant6_beta_div_neutral_met$comparison <- factor(
  transplant6_beta_div_neutral_met$comparison,
  levels = c("Transplant - 1_Acclimation",  # Acclimation-Transplant first
             "6_Post-FMT2 - Transplant",    # Transplant-Treatment second
             "6_Post-FMT2 - 1_Acclimation")    # Treatment-Acclimation last
)

transplant6_beta_div_neutral_met_filtered <- transplant6_beta_div_neutral_met %>%
  filter(!(Sample_1_part == "Transplant" & Sample_2_part == "Transplant"))
```

```{r diss_plot2, echo=FALSE, results=FALSE}
Acclimation_Transplant<-transplant6_beta_div_neutral_met_filtered %>% 
  filter(comparison == "Transplant - 1_Acclimation")
stat.test_accli <- Acclimation_Transplant %>% 
  t_test(Distance ~ type)%>% 
  add_xy_position(x = "type")%>% 
  filter(p.adj < 0.05)


bxp1 <- Acclimation_Transplant %>%
  ggplot(aes(x = type, y = Distance, color = type, fill = type)) +
  geom_boxplot() +
  scale_color_manual(name = "Type",
                     breaks = c("Control", "Hot_control", "Treatment"),
                     labels = c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                     values = c("#4477AA", "#d57d2c", "#76b183")) +
  scale_fill_manual(name = "Type",
                    breaks = c("Control", "Hot_control", "Treatment"),
                    labels = c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                    values = c("#4477AA50", "#d57d2c50", "#76b18350")) +
  scale_x_discrete(labels = c("Control" = "Cold-Cold", "Hot_control" = "Hot-Hot", "Treatment" = "Cold-Hot")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Beta diversity dissimilarity distance")

# Add significance bars with p-values
p1<-bxp1 + 
  geom_signif(comparisons = list(c("Control", "Hot_control"), c("Control", "Treatment"), c("Hot_control", "Treatment")),
              map_signif_level = TRUE)


Transplant_Treatment<-transplant6_beta_div_neutral_met_filtered %>% 
  filter(comparison == "6_Post-FMT2 - Transplant")
bxp2 <- Transplant_Treatment %>%
  ggplot(aes(x = type, y = Distance, color = type, fill = type)) +
  geom_boxplot() +
  scale_color_manual(name = "Type",
                     breaks = c("Control", "Hot_control", "Treatment"),
                     labels = c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                     values = c("#4477AA", "#d57d2c", "#76b183")) +
  scale_fill_manual(name = "Type",
                    breaks = c("Control", "Hot_control", "Treatment"),
                    labels = c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                    values = c("#4477AA50", "#d57d2c50", "#76b18350")) +
  scale_x_discrete(labels = c("Control" = "Cold-Cold", "Hot_control" = "Hot-Hot", "Treatment" = "Cold-Hot")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Beta diversity dissimilarity distance")
# Add significance bars with p-values
p2<-bxp2 + 
  geom_signif(comparisons = list(c("Control", "Hot_control"), c("Control", "Treatment"), c("Hot_control", "Treatment")),
              map_signif_level = TRUE)

Treatment_Acclimation<-transplant6_beta_div_neutral_met_filtered %>% 
  filter(comparison == "6_Post-FMT2 - 1_Acclimation")
bxp3 <- Treatment_Acclimation %>%
  ggplot(aes(x = type, y = Distance, color = type, fill = type)) +
  geom_boxplot() +
  scale_color_manual(name = "Type",
                     breaks = c("Control", "Hot_control", "Treatment"),
                     labels = c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                     values = c("#4477AA", "#d57d2c", "#76b183")) +
  scale_fill_manual(name = "Type",
                    breaks = c("Control", "Hot_control", "Treatment"),
                    labels = c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                    values = c("#4477AA50", "#d57d2c50", "#76b18350")) +
  scale_x_discrete(labels = c("Control" = "Cold-Cold", "Hot_control" = "Hot-Hot", "Treatment" = "Cold-Hot")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("Beta diversity dissimilarity distance")

# Add significance bars with p-values
p3<-bxp3 + 
  geom_signif(comparisons = list(c("Control", "Hot_control"), c("Control", "Treatment"), c("Hot_control", "Treatment")),
              map_signif_level = TRUE)
```

```{r diss_plot3, , fig.height=7, eho=FALSE, results=TRUE}
ggarrange( p1, p2, p3, ncol=3, nrow=1, common.legend = TRUE, legend="right")
```


### 6. Are there differences between the control and the treatment group?

#### After 1 week --> Post-FMT1

```{r beta_div_phylo_post1, echo=TRUE, results=FALSE}
post1 <- meta %>%
  filter(time_point == "5_Post-FMT1")

post1.counts <- temp_genome_counts[,which(colnames(temp_genome_counts) %in% rownames(post1))]
identical(sort(colnames(post1.counts)),sort(as.character(rownames(post1))))

post1_nmds <- sample_metadata %>%
  filter(time_point == "5_Post-FMT1")
```

#### Number of samples used

```{r beta_div_neutral_treat1_samples_2, echo=FALSE, results = TRUE, message=FALSE, warning=FALSE, comments=FALSE}
nrow(post1)
```

```{r beta_div_phylo_post1_1, echo=TRUE, results = FALSE, message=FALSE, warning=FALSE, comments=FALSE, eval=FALSE}
beta_div_richness_post1<-hillpair(data=post1.counts, q=0)
beta_div_neutral_post1<-hillpair(data=post1.counts, q=1)
beta_div_phylo_post1<-hillpair(data=post1.counts, q=1, tree=genome_tree)
beta_div_func_post1<-hillpair(data=post1.counts, q=1, dist=dist)
```

```{r beta_div_phylo_post1_2, echo=TRUE, results = FALSE, message=FALSE, warning=FALSE, comments=FALSE}
#Arrange of metadata dataframe
post1_arrange<-post1[labels(beta_div_neutral_post1$S),]
```

##### Richness

```{r beta_richness_permanova_post1_2, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_richness_post1$S, post1$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_richness_permanova_post1_1, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_richness_post1$S ~ species+type, data=post1[labels(beta_div_richness_post1$S),], permutations=999) %>%
		as.matrix() %>%
		kable()
```

```{r beta_richness_permanova_post1_3, echo=FALSE, warning=FALSE, comments="", message=FALSE}
pairwise<-pairwise.adonis(beta_div_richness_post1$S,post1_arrange$type, perm=999)
pairwise
```

##### Neutral

```{r beta_neutral_permanova_post1_2, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_neutral_post1$S, post1$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_neutral_permanova_post1_1, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_neutral_post1$S ~ species+type, data=post1[labels(beta_div_neutral_post1$S),], permutations=999) %>%
		as.matrix() %>%
		kable()
```

```{r beta_neutral_permanova_post1_3, echo=FALSE, warning=FALSE, comments="", message=FALSE}
pairwise<-pairwise.adonis(beta_div_neutral_post1$S,post1_arrange$type, perm=999)
pairwise
```

##### Phylogenetic

```{r beta_phylo_permanova_post1_2, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_phylo_post1$S, post1$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_phylo_permanova_post1_1, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_phylo_post1$S ~ species+type, data=post1[labels(beta_div_phylo_post1$S),], permutations=999) %>%
		as.matrix() %>%
		kable()
```

```{r beta_phylo_permanova_post1_3, echo=FALSE, warning=FALSE, comments="", message=FALSE}
pairwise<-pairwise.adonis(beta_div_phylo_post1$S,post1_arrange$type, perm=999)
pairwise
```

##### Functional

```{r beta_func_permanova_post1_2, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_func_post1$S, post1$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_func_permanova_post1_1, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_func_post1$S ~ species+type, data=post1[labels(beta_div_func_post1$S),], permutations=999) %>%
		as.matrix() %>%
		kable()
```

```{r beta_func_permanova_post1_3, echo=FALSE, warning=FALSE, comments="", message=FALSE}
pairwise<-pairwise.adonis(beta_div_func_post1$S,post1_arrange$type, perm=999)
pairwise
```


```{r beta_neutral_nmds_post1, echo=TRUE, warning=FALSE, comments="", message=FALSE, results="hide"}
beta_richness_nmds_post1 <- beta_div_richness_post1$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(post1_nmds, by = join_by(sample == Tube_code))

beta_neutral_nmds_post1 <- beta_div_neutral_post1$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(post1_nmds, by = join_by(sample == Tube_code))

beta_phylogenetic_nmds_post1 <- beta_div_phylo_post1$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(post1_nmds, by = join_by(sample == Tube_code))

beta_functional_nmds_post1 <- beta_div_func_post1$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(post1_nmds, by = join_by(sample == Tube_code))
```

```{r beta_neutral_nmds_post1_plot, eho=FALSE, results=FALSE}
p0<-beta_richness_nmds_post1 %>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
        scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Richness beta diversity") +
				theme_classic() +
				theme(legend.position="none")

p1<-beta_neutral_nmds_post1 %>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
        scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Neutral beta diversity") +
				theme_classic() +
				theme(legend.position="none")
  
p2<-beta_phylogenetic_nmds_post1 %>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
				scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (), x="NMDS1 \n Phylogenetic beta diversity") +
				theme_classic() +
				theme(legend.position="none")

p3<-beta_functional_nmds_post1 %>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
				scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (), x="NMDS1 \n Functional beta diversity") +
				theme_classic()+
				theme(legend.position="none")
```

```{r beta_neutral_nmds_plot_post1_final, fig.height=7, eho=FALSE, results=TRUE}
ggarrange(p0, p1, p2, p3, ncol=2, nrow=2, common.legend = TRUE, legend="right")
```


#### After 2 weeks -->Post-FMT2

```{r beta_div_phylo_post2, echo=TRUE, results=FALSE}
post2 <- meta %>%
  filter(time_point == "6_Post-FMT2")

post2.counts <- temp_genome_counts[,which(colnames(temp_genome_counts) %in% rownames(post2))]
identical(sort(colnames(post2.counts)),sort(as.character(rownames(post2))))

post2_nmds <- sample_metadata %>%
  filter(time_point == "6_Post-FMT2")
```

#### Number of samples used

```{r beta_div_neutral_treat_samples_2, echo=FALSE, results = TRUE, message=FALSE, warning=FALSE, comments=FALSE}
nrow(post2)
```

```{r beta_div_phylo_post2_1, echo=TRUE, results = FALSE, message=FALSE, warning=FALSE, comments=FALSE, eval=FALSE}
beta_div_richness_post2<-hillpair(data=post2.counts, q=0)
beta_div_neutral_post2<-hillpair(data=post2.counts, q=1)
beta_div_phylo_post2<-hillpair(data=post2.counts, q=1, tree=genome_tree)
beta_div_func_post2<-hillpair(data=post2.counts, q=1, dist=dist)
```

```{r beta_div_phylo_post2_2, echo=TRUE, results = FALSE, message=FALSE, warning=FALSE, comments=FALSE}
#Arrange of metadata dataframe
post2_arrange<-post2[labels(beta_div_neutral_post2$S),]
```

##### Richness

```{r beta_richness_permanova_post2_2, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_richness_post2$S, post2$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_richness_permanova_post2_1, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_richness_post2$S ~ type, data=post2[labels(beta_div_richness_post2$S),], permutations=999) %>%
		as.matrix() %>%
		kable()
```

```{r beta_richness_permanova_post2_3, echo=FALSE, warning=FALSE, comments="", message=FALSE}
pairwise<-pairwise.adonis(beta_div_richness_post2$S,post2_arrange$type, perm=999)
pairwise
```

##### Neutral

```{r beta_neutral_permanova_post2_2, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_neutral_post2$S, post2$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_neutral_permanova_post2_1, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_neutral_post2$S ~ type, data=post2[labels(beta_div_neutral_post2$S),], permutations=999) %>%
		as.matrix() %>%
		kable()
```

```{r beta_neutral_permanova_post2_3, echo=FALSE, warning=FALSE, comments="", message=FALSE}
pairwise<-pairwise.adonis(beta_div_neutral_post2$S,post2_arrange$type, perm=999)
pairwise
```

##### Phylogenetic

```{r beta_phylo_permanova_post2_3, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_phylo_post2$S, post2$type) %>% permutest(., pairwise = TRUE)
```


```{r beta_phylo_permanova_post2_1, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_phylo_post2$S ~ type, data=post2[labels(beta_div_phylo_post2$S),], permutations=999) %>%
		as.matrix() %>%
		kable()
```

```{r beta_phylo_permanova_post2_2, echo=FALSE, warning=FALSE, comments="", message=FALSE}
pairwise<-pairwise.adonis(beta_div_phylo_post2$S,post2_arrange$type, perm=999)
pairwise
```

##### Functional

```{r beta_func_permanova_post2_2, echo=FALSE, warning=FALSE, comments="", message=FALSE}
betadisper(beta_div_func_post2$S, post2$type) %>% permutest(., pairwise = TRUE)
```

```{r beta_func_permanova_post2_1, echo=FALSE, warning=FALSE, comments="", message=FALSE}
adonis2(formula=beta_div_func_post2$S ~ type, data=post2[labels(beta_div_func_post2$S),], permutations=999) %>%
		as.matrix() %>%
		kable()
```

```{r beta_func_permanova_post2_3, echo=FALSE, warning=FALSE, comments="", message=FALSE}
pairwise<-pairwise.adonis(beta_div_func_post2$S,post2_arrange$type, perm=999)
pairwise
```


```{r beta_neutral_nmds_post2, echo=TRUE, warning=FALSE, comments="", message=FALSE, results="hide"}
beta_richness_nmds_post2 <- beta_div_richness_post2$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(post2_nmds, by = join_by(sample == Tube_code))

beta_neutral_nmds_post2 <- beta_div_neutral_post2$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(post2_nmds, by = join_by(sample == Tube_code))

beta_phylogenetic_nmds_post2 <- beta_div_phylo_post2$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(post2_nmds, by = join_by(sample == Tube_code))

beta_functional_nmds_post2 <- beta_div_func_post2$S %>%
				metaMDS(.,trymax = 500, k=2, verbosity=FALSE) %>%
				scores() %>%
				as_tibble(., rownames = "sample") %>%
				left_join(post2_nmds, by = join_by(sample == Tube_code))
```


```{r beta_neutral_nmds_post2_plot, eho=FALSE, results=FALSE}
p0<-beta_richness_nmds_post2 %>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
        scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Richness beta diversity") +
				theme_classic() +
				theme(legend.position="none")

p1<-beta_neutral_nmds_post2 %>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
        scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y = "NMDS2", x="NMDS1 \n Neutral beta diversity") +
				theme_classic() +
				theme(legend.position="none")
  
p2<-beta_phylogenetic_nmds_post2 %>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
				scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (), x="NMDS1 \n Phylogenetic beta diversity") +
				theme_classic() +
				theme(legend.position="none")

p3<-beta_functional_nmds_post2 %>%
			group_by(type) %>%
			mutate(x_cen = mean(NMDS1, na.rm = TRUE)) %>%
			mutate(y_cen = mean(NMDS2, na.rm = TRUE)) %>%
			ungroup() %>%
			ggplot(., aes(x=NMDS1,y=NMDS2, color=type)) +
				scale_color_manual(name="Type",
                       breaks=c("Control", "Hot_control", "Treatment"),
                       labels=c("Cold-Cold", "Hot-Hot", "Cold-Hot"),
                       values=c("#4477AA","#d57d2c","#76b183")) +
				geom_point(size=2) +
				geom_segment(aes(x=x_cen, y=y_cen, xend=NMDS1, yend=NMDS2), alpha=0.2) +
        labs(y= element_blank (), x="NMDS1 \n Functional beta diversity") +
				theme_classic()+
				theme(legend.position="none")
```

```{r beta_neutral_nmds_plot_post2_final, fig.height=7, eho=FALSE, results=TRUE}
ggarrange(p0, p1, p2, p3, ncol=2, nrow=2, common.legend = TRUE, legend="right")
```
