# Physiological measurements

## Respirometry

```{r load_data_resp}
load("data/data_27022025.Rdata")
respirometry_resp2 <- read_excel("data/respirometry_resp2.xlsx")
```

```{r dataframe_modify, warning=FALSE}
respirometry_resp2$individual <- as.factor(respirometry_resp2$individual)
respirometry_resp2$cage <- as.factor(respirometry_resp2$cage)
respirometry_resp2$time_point <- as.factor(respirometry_resp2$time_point)
respirometry_resp2$type <- as.factor(respirometry_resp2$type)
respirometry_resp2$population <- as.factor(respirometry_resp2$population)
respirometry_resp2$QC_normalized <- as.numeric(respirometry_resp2$QC_normalized)
respirometry_resp2$r2 <- as.numeric(respirometry_resp2$r2)


#Filter NA values
respirometry_resp2<-respirometry_resp2 %>%
  filter(QC_normalized!="NA")
```

### Acclimation oxygen consumption

```{r plot}
respirometry_resp2 %>%
  filter(time_point=="0")%>%
  filter(individual!="AH1_2nd_8") %>%
  filter(individual!="LI1_2nd_4") %>%
  filter(individual!="LI1_2nd_3") %>% #lizard doesn't stop moving during the whole measurement, increasing the value and making it and outlier
  filter(individual!="LI1_2nd_10") %>%
  ggplot(aes(x = population, y = QC_normalized, color=population, fill=population, alpha=0.2)) +
    geom_boxplot()+
    geom_jitter() +
      scale_color_manual(name="Population",
          breaks=c("Cold","Warm"),
          labels=c("Cold","Warm"),
          values=c('#008080', "#d57d2c")) +
      scale_fill_manual(name="Population",
          breaks=c("Cold","Warm"),
          labels=c("Cold","Warm"),
          values=c('#00808050', "#d57d2c50"))+
    #geom_text(aes(label = individual), hjust = 0, nudge_x = 0.05) +
    theme(axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))+
    stat_compare_means(size=3,vjust = 3, hjust=1 )+
    theme(legend.position="none",
          panel.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())+
    theme_minimal()+
    labs(x = "Population", y= "Normalized O2 consumption (ml/s/g)")
```

### Time-point oxygen consumption

```{r plot_1}
respirometry_resp2 %>%
  filter(weight!=5.19)%>%
  filter(weight!=5.91)%>%
  filter(weight!=5.03)%>%
  filter(r2!=0.808)%>%
  filter(QC_normalized!=0.000036) %>% #r2 low
  filter(r2!=0.728) %>%  #r2 low
  ggplot(aes(x = time_point, y = QC_normalized, color=time_point, fill=time_point, alpha=0.2, label=cage)) +
    geom_boxplot()+
    geom_jitter() +
      scale_color_manual(name="Time_point",
          breaks=c("0","1","2"),
          labels=c("Acclimation","1 week post FMT", "2 weeks post FMT"),
          values=c('#BFA366', "#dec14b", "#6E5244")) +
      scale_fill_manual(name="Time_point",
          breaks=c("0","1","2"),
          labels=c("Acclimation","1 week post FMT", "2 weeks post FMT"),
          values=c('#BFA36650', "#dec14b50","#6E524450"))+
    geom_text()+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
    facet_wrap(~ factor(type))+
    stat_compare_means(size=3)+
    theme(legend.position="none")+
    labs(x = "Time_point", y= "Normalized O2 consumption (ml/s/g)")
```

```{r plot_1_1}
respirometry_resp2 %>%
  filter(weight!=5.19)%>%
  filter(weight!=5.91)%>%
  filter(weight!=5.03)%>%
  filter(r2!=0.808)%>%
  filter(QC_normalized!=0.000036) %>% #r2 low
  filter(r2!=0.728) %>%
  ggplot(aes(x = time_point, y = QC_normalized, color=time_point, fill=time_point, alpha=0.2)) +
    geom_boxplot()+
    geom_jitter() +
    scale_color_manual(name="Time_point",
          breaks=c("0","1","2"),
          labels=c("Acclimation","1 week post FMT", "2 weeks post FMT"),
          values=c('#BFA366', "#dec14b", "#6E5244")) +
      scale_fill_manual(name="Time_point",
          breaks=c("0","1","2"),
          labels=c("Acclimation","1 week post FMT", "2 weeks post FMT"),
          values=c('#BFA36650', "#dec14b50","#6E524450"))+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
    facet_wrap(~ factor(type))+
    stat_compare_means(size=3)+
    theme(legend.position="none")+
    labs(x = "Time_point", y= "Normalized O2 consumption (ml/s/g)")+
    geom_signif(comparisons = list(c("0", "1"), c("0", "2"), c("1", "2")),
              map_signif_level = TRUE)
```

### Data modelling 

#### CC vs CI Raw data modelling (LMM)

```{r model, eval=TRUE}
#subset all the wrong measurements
respirometry_resp2_subset<-respirometry_resp2%>%
  filter(type!="WC")%>%
  filter(weight!=5.19)%>%
  filter(weight!=5.91)%>%
  filter(weight!=5.03)%>%
  filter(r2!=0.808)%>%
  filter(QC_normalized!=0.000036) %>% #r2 low
  filter(r2!=0.728) %>%
  filter(cage!="Cold_17") %>%
  filter(cage!="Cold_3") %>%
  filter(cage!="Cold_6") %>%
  filter(cage!="Cold_9") %>%
  filter(cage!="Hot_2") %>%
  filter(cage!="Hot_8")

#Create linear model formula
model <- lme(fixed = QC_normalized ~  time_point+type, data = respirometry_resp2_subset,
             random = ~ 1 | individual)

#Print the model summary
summary(model)
MuMIn::r.squaredGLMM(model)
```


#### CI vs WC Raw data modelling (LMM)

```{r model_2, eval=TRUE}
#Filter out WC from the dataframe and create a subset
respirometry_resp2_subset<-respirometry_resp2%>%
  filter(weight!=5.19)%>%
  filter(weight!=5.91)%>%
  filter(weight!=5.03)%>%
  filter(r2!=0.808)%>%
  filter(QC_normalized!=0.000036) %>% #r2 low
  filter(r2!=0.728) %>%
  filter(cage!="Cold_17") %>%
  filter(cage!="Cold_3") %>%
  filter(cage!="Cold_6") %>%
  filter(cage!="Cold_9") %>%
  filter(cage!="Hot_2") %>%
  filter(cage!="Hot_8")

#Create linear model formula
model <- lme(fixed = QC_normalized ~  time_point+type, data = respirometry_resp2_subset,
             random = ~ 1 | individual)

#Print the model summary
summary(model)
MuMIn::r.squaredGLMM(model)
```

## PHA test

```{r load_data_pha, warning=FALSE, comment=FALSE}
pha_rep2 <- read_tsv("data/Fito_2023.tsv")
```

```{r dataframe_modify_2}
pha_rep2$Individual<-as.factor(pha_rep2$Individual)
pha_rep2$Cage<-as.factor(pha_rep2$Cage)
pha_rep2$time_point<-as.factor(pha_rep2$time_point)
pha_rep2$Type<-as.factor(pha_rep2$Type)
```

### PHA measurements

```{r plot_pha}
pha_rep2 %>%
  ggplot(aes(x = Weight, y = Mean_value)) +
    geom_point() +
    geom_smooth(method = lm, formula = y ~ x) +
        ggpmisc::stat_poly_eq()+
  facet_nested(~ time_point+Type)+
    labs(x = "Weight (g)", y="Measurement (mm)")
```

#### Average PHA measurements

```{r pha_measurement}
pha_rep2 %>%
  rowwise() %>%
  mutate(
    mean = round(mean(c(Value1, Value2, Value3, Value4, Value5)), 2),
    sd = round(sd(c(Value1, Value2, Value3, Value4, Value5)), 2)
  ) %>%
  unite("Average", mean, sd, sep = " ± ", remove = TRUE) %>%
  select(Individual, time_point, Average) %>%
    tt()
```

### Time-point measurements

```{r plot_pha_5}
pha_rep2 %>%
  ggplot(aes(x = time_point, y = normalized, color=time_point, fill=time_point, alpha=0.2 )) +
  geom_boxplot()+
  geom_jitter() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_color_manual(name="Time_point",
          breaks=c("0","1"),
          labels=c("Pre","24h"),
          values=c('#BFA366', "#dec14b")) +
      scale_fill_manual(name="Time_point",
          breaks=c("0","1"),
          labels=c("Pre","24h"),
          values=c('#BFA36650', "#dec14b50"))+
  facet_wrap(~ factor(Type))+
  stat_compare_means(size=3)+
  theme(legend.position="none")+
  labs(x = "Time_point", y= "Normalized measurement (mm/g)")
```

### Time-point measurements, linking individuals

```{r plot_pha_1, warning=FALSE}
pha_rep2 %>%
  ggplot(aes(x = interaction(time_point, Individual), y = normalized, color=time_point, fill=time_point, alpha=0.2 )) +
  geom_boxplot()+
  geom_point(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8), alpha = 0.5) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  geom_line(aes(group=Individual)) +
  scale_color_manual(name="Time_point",
          breaks=c("0","1"),
          labels=c("Pre","24h"),
          values=c('#BFA366', "#dec14b")) +
      scale_fill_manual(name="Time_point",
          breaks=c("0","1"),
          labels=c("Pre","24h"),
          values=c('#BFA36650', "#dec14b50"))+
  facet_wrap(~ factor(Type),scales = "free")+
  theme(legend.position="none")+
  labs(x = "Time_point", y= "Normalized measurement (mm/g)")
```


### Data modelling

#### Raw data modelling (LMM)

```{r model_pha}
#Create linear model formula
model6 <- lme(fixed = normalized ~  time_point*Type, data = pha_rep2,
               random = ~ 1 | Individual)

#Print the model summary
summary(model6)
MuMIn::r.squaredGLMM(model6)
```

#### Raw data modelling (LM) without random effect

```{r model_pha_1}
#Create linear model formula
formula7 <- formula(normalized ~ time_point*Type)

#Fit the linear model
model7 <- lm(formula7, data = pha_rep2)

#Print the model summary
summary(model7)
```

### Data modeling without WC

#### Raw data modelling (LMM)

```{r model_pha_2}
#Filter out WC measurements
pha_rep2_cold<-pha_rep2 %>%
  filter(Type!="WC")

#Create linear model formula
model8 <- lme(fixed = normalized ~  time_point*Type, data = pha_rep2_cold,
               random = ~ 1 | Individual)

#Print the model summary
summary(model8)
MuMIn::r.squaredGLMM(model8)
```

#### Raw data modelling (LM) without random effect

```{r model_pha2_1}
#Create linear model formula
formula9 <- formula(normalized ~ time_point*Type)

#Fit the linear model
model9 <- lm(formula9, data = pha_rep2_cold)

#Print the model summary
summary(model9)
```